### 2023-02-26 appendixA.tex

\newpage
\section{Appendix A: An implementation of union-findable Possible Type Sets}

\begin{figure}[h!]
\begin{lstlisting}[escapeinside={(*}{*)}]
open UnionFind

type t = possible_types UnionFind.elem
and possible_type_set = possible_type list
and possible_type =
  | Unknown(provenance)
  | Num
  | Bool
  | Arrow of t * t;

let rec merge_sets_UF (s1: t, s2: t): t =
    let merged_typs = merge_sets (UnionFind.get s1) (UnionFind.get s2) in
    let final_rep = UnionFind.union s1 s2 in
    UnionFind.set final_rep merged_typs;
    final_rep
and merge_sets (s1: possible_type_set, s2: possible_type_set): possible_types = 
    match (s2)
    | [] -> s1
    | s2_hd::s2_hd -> merge_sets (add s1 s2_hd) s2_hd
and add (s: possible_type_set, t: possible_type): possible_type_set =
    match (s)
    | [] -> [t]
    | s_hd::s_tl -> (
        match (s_hd, t)
        | (Arrow(s_hd_L, s_hd_R), Arrow(t_L, t_R)) ->
            let left = merge_sets_UF s_hd_L t_L in
            let right = merge_sets_UF s_hd_R t_R in
            Arrow(left, right)::s_tl
        | _ -> if (s_hd = t) then s else s_hd::(add s_tl t)
    )

let union => if (UnionFind.eq t1 t2) then () else merge_sets_UF t1 t2;
\end{lstlisting}
\vspace{-4px}
 \hrule
\caption{An Implementation of \textit{PossibleTypeSet} using UnionFind}
\label{fig:ufalgcode}
\end{figure}