### 2023-03-02 supplement/unionFindPTS.tex

\newpage
\section{An implementation of union-findable Possible Type Sets}

\begin{figure}[h!]
\begin{lstlisting}[escapeinside={(*}{*)}]
open UnionFind

type t = possible_types UnionFind.elem
and possible_type_set = possible_type list
and possible_type =
  | Unknown(provenance)
  | Num
  | Bool
  | Arrow of t * t;

let rec merge_sets_UF (s1: t, s2: t): t =
    let merged_typs = merge_sets (UnionFind.get s1) (UnionFind.get s2) in
    let final_rep = UnionFind.union s1 s2 in
    UnionFind.set final_rep merged_typs;
    final_rep
and merge_sets (s1: possible_type_set, s2: possible_type_set): possible_types = 
    match (s2)
    | [] -> s1
    | s2_hd::s2_hd -> merge_sets (add s1 s2_hd) s2_hd
and add (s: possible_type_set, t: possible_type): possible_type_set =
    match (s)
    | [] -> [t]
    | s_hd::s_tl -> (
        match (s_hd, t)
        | (Arrow(s_hd_L, s_hd_R), Arrow(t_L, t_R)) ->
            let left = merge_sets_UF s_hd_L t_L in
            let right = merge_sets_UF s_hd_R t_R in
            Arrow(left, right)::s_tl
        | _ -> if (s_hd = t) then s else s_hd::(add s_tl t)
    )

let (*$\amalg_{UF}$*) => if (UnionFind.eq t1 t2) then () else merge_sets_UF t1 t2;

\end{lstlisting}
 \hrule
\caption{An Implementation of \textit{PossibleTypeSet} using UnionFind}
\label{fig:ufalgcode}
\end{figure}

\begin{figure}[htb!]
\begin{lstlisting}[escapeinside={(*}{*)}]
let create_traversable_edge (graph, t1, t2) =
    let (rep1, rep2) = (UnionFind.find(t1), UnionFind.find(t2)) in
    let (s1, s2) = (Map.find(graph, rep1), Map.find(graph, rep2)) in
    let union_rep = UnionFind.union(rep1, rep2) in
    let s_merged = s1 (*$\amalg_{UF}$*) s2 in
    Map.update(graph, union_rep, s_merged);;
let create_solution_edge (graph, t1, t2) =
    let rep1 = UnionFind.find(t1) in
    let s = Map.find(graph, rep1) in
    let s' = s (*$\amalg_{UF}$*) UnionFind.elem(single(t2)) in
    Map.update(graph, t1, s');;
\end{lstlisting}
\hrule
\caption{Constraint solving algorithm helper functions, using Union-Find.}
\label{fig:algcode_construct_graph_helpers}
\end{figure}

\begin{figure}[htb!]
\begin{lstlisting}[escapeinside={(*}{*)}]
let handle_occurs_check (graph) =
    let annotate_with_error (key) =
        let s = Map.find(graph, UnionFind.find(key) in
        let pass_occurs = is_acyclic(s)
        (key, s, pass_occurs)
    in
    List.map(Map.keys(graph), annotate_with_error);;
\end{lstlisting}
 \hrule
\caption{Occurs check assessment}
\label{fig:occurs_check}
\end{figure}