### 2023-03-02 calculus.tex
Some edits on section 2

\section{The Marked Lambda Calculus}
\label{sec:calculus}

Programmers often work with program states that are syntactically well-formed but not well-typed.
\Cref{fig:calculus-examples} shows common type errors as they appear in the Hazel programming
environment. For example, a simple error that may occur is the use of an unbound variable. As in
\cref{fig:calculus-examples-unbound}, the editor indicates that the usage of an unbound $y$ is
invalid by highlighting it in red.

\input{figs/calculus-examples}

Sometimes, the type an expression \emph{should} be is known. Consider
\cref{fig:calculus-examples-inconsistent-types}, in which it is expected that both operands of the
$+$ operator are numbers. Since $\textsf{true}$ is not a number, a type error arises, and Hazel
indicates a mismatch between the expected and actual types of the operand. The application in
\cref{fig:calculus-examples-app-non-lambda} of a non-lambda leads to a similar kind of error. In
this case, however, the expectation is not a singular type, but any member of a family of function
types. \Cref{fig:calculus-examples-inconsistent-branches} presents an error of inconsistent
branches: the branches of the conditional are of different types: $5$ is a number, while
$\textsf{false}$ is a boolean. Hazel opts to highlight the entire expression in such a situation,
while other tools may choose one branch to be ``correct'' and ``blame'' the type mismatch error on
the other.

% TODO elaborate on this part

Unfortunately, formal typing semantics usually only give meaning to well-typed programs, leaving
such ill-typed ones meaningless. This is insufficient for editor services and language servers,
which must provide users feedback on where and why a program is ill-typed. Moreover, they must
continue providing semantic services even when an error is encountered. Indeed, similarly to
error-recovering parsers, real-world type checkers must robustly recover from type errors.
Traditionally, this has been performed in an \emph{ad hoc} manner [citations?], with different
languages and tools adopting varying approaches in the localization, reportage, and recovery of
errors.

The key contribution of this section is the \emph{marked lambda calculus}, a formal calculus for
bidirectional type error localization. Namely, the calculus employs bidirectional and gradual typing
to provide a judgmental framework with a unified theory for the design of typing semantics that
support ill-typed programs. \Cref{sec:calculus-calculus} describes its form and metatheory by
example on a small gradually typed lambda calculus (GTLC) \cite{Siek06a}, extending it with binary
product types in \cref{sec:calculus-products}. All rules and theorems may be found in the
supplementary appendix, alongside a complete mechanization in the Agda proof assistant
\cite{norell:thesis} (discussed in \cref{sec:calculus-agda}). With products, we then consider how
error localization is complicated by patterned let expressions in \cref{sec:calculus-let}. Finally,
\cref{sec:calculus-poly} explores how the system might be extended to more complex typing features,
such as System F-style polymorphism.

% TODO not really sure when to start talking about bidirectional typing

% \emph{Bidirectional typing} combines type checking, which determines if a program satisfies a given
% type, and type synthesis, which generates a type from the program. This gives a simple algorithmic
% system in which type information is recursively propagated through terms and allows for local type
% inference.

% TODO talk about idea that bidirectional typing makes error localization easy? i.e. why choose
% bidirectional?

\subsection{The Calculus}
\label{sec:calculus-calculus}

% TODO outline the structure of this section at the beginning
% TODO make it clear unmarked/marked languages share same types

We now formally introduce the marked lambda calculus through its application to the GTLC extended
with numbers and booleans. Consider the syntax for such a language, given as $\EMName$ in
\cref{fig:calculus-syntax}. The base types $\TNum$ and $\TBool$ classify number and boolean
expressions. The number literal corresponding to the mathematical number $n$ is given by $\ENumMV$,
and there is a single arithmetic operation of addition on numeric expressions. $\ETrue$ and
$\EFalse$ correspond to the boolean values $\textsf{true}$ and $\textsf{false}$, respectively, and
$\EIf{\EMV_1}{\EMV_2}{\EMV_3}$ gives a conditional structure.

In addition to these forms, $\TUnknown$ is the unknown type from gradual type theory \cite{Siek06a}.
Modelling the edit state of a program in development, $\EEHole$ denotes an \emph{expression hole},
used to represent syntactically incomplete portions of the program à la Hazelnut
\cite{HazelnutPOPL}. Note, however, that these are not critical to the calculus.

\begin{figure}[htbp]
  \[\begin{array}{rrcl}
    \TMName  & \TMV  & \Coloneqq & \TUnknown \mid \TNum \mid \TBool \mid \TArrow{\TMV}{\TMV} \mid \TProd{\TMV}{\TMV} \\
    \EMName  & \EMV  & \Coloneqq & x \mid \ELam{x}{\TMV}{\EMV} \mid \EAp{\EMV}{\EMV} \mid \ELet{x}{\EMV}{\EMV}
                       \mid           \ENumMV \mid \EPlus{\EMV}{\EMV} \\
             &       & \mid         & \ETrue \mid \EFalse \mid \EIf{\EMV}{\EMV}{\EMV}
                       \mid           \EPair{\EMV}{\EMV}
                       \mid           \EProjL{\EMV} \mid \EProjR{\EMV}
                       \mid           \EEHole \\
    \ECMName & \ECMV & \Coloneqq & x \mid \ECLam{x}{\TMV}{\ECMV} \mid \ECAp{\ECMV}{\ECMV} \mid \ECLet{x}{\ECMV}{\ECMV}
                       \mid           \ECNumMV \mid \ECPlus{\ECMV}{\ECMV} \\
             &       & \mid         & \ECTrue \mid \ECFalse \mid \ECIf{\ECMV}{\ECMV}{\ECMV}
                       \mid           \ECPair{\ECMV}{\ECMV} \mid \ECProjL{\ECMV} \mid \ECProjR{\ECMV}
                       \mid           \ECEHole \\
             &       & \mid         & \ECUnbound{x} \mid \ECInconType{\ECMV} \mid \ECInconBr{\ECMV}{\ECMV}{\ECMV} \mid \ECInconAsc{\ECMV} \\
             &       & \mid         & \ECSynNonMatchedArrow{\ECMV} \mid \ECAnaNonMatchedArrow{\ECMV}
                       \mid           \ECSynNonMatchedProd{\ECMV} \mid \ECAnaNonMatchedProd{\ECMV}
  \end{array}\]
  \caption{Syntax of an extended GTLC and its marked version.}
  \label{fig:calculus-syntax}
\end{figure}

Types classify expressions by a \emph{bidirectional type system} \cite{BidirTyping}, which employs
two mutually defined judgments. \emph{Type synthesis}, written $\ctxSynTypeU{\ctx}{\EMV}{\TMV}$,
establishes that, under the typing context $\ctx$, the expression $e$ synthesizes or locally infers
\cite{Localinf} the type $\TMV$. \emph{Type analysis}, written $\ctxAnaTypeU{\ctx}{\EMV}{\TMV}$,
states that the expression $\EMV$ may appear where an expression of type $\TMV$ is expected. We may
give these typing semantics traditionally, without consideration for error localization and
recovery.

The marked lambda calculus builds atop these typing semantics into a two-layer system:
%
\begin{itemize}
  \item The \textbf{unmarked language}, which is the original language. Expressions of this language
    are called \emph{unmarked expressions}, denoted by the metavariable $\EMV$.

  \item The \textbf{marked language}, which mirrors the structure of the unmarked language but is
    extended with explicit \emph{error marks}. We call expressions of this language to be
    \emph{marked expressions}, denoted $\ECMV$.
    % TODO more intuition for what marked expressions are?

  \item \textbf{Marking}, which \emph{marks} any unmarked expression into a marked expression,
    inserting error marks where appropriate.
\end{itemize}
%
In other words, we extend the ordinary typing semantics of a language with a secondary ``marked''
language and ``mark'' unmarked programs into marked ones. This corresponds to a type checking
process with error localization and recovery. The marked language may then serve as a foundation for
other semantic services. % TODO such as THI.

\Cref{fig:calculus-syntax} furthermore gives the syntax for marked expressions (as $\ECMName$),
which replicates all the forms of unmarked expressions. There are additionally several kinds of
\emph{error marks}, which represent the different type errors that may arise. Intuitively, they each
denote a different error message that might be shown by an editor or emitted by a compiler.

Throughout the remainder of this section, as we formulate marking for the GTLC, we will motivate and
give precise semantics for these error marks. Note that these constitute a non-exhaustive list;
depending on the syntactic forms and typing features found in a given language, other kinds of marks
may be necessary. Indeed, in \cref{sec:calculus-poly}, we consider the kinds of marks that may be
required in an language with explicit polymorphism.

The marked language possesses its own typing semantics, also formulated bidirectionally. We write
$\ctxSynTypeM{\ctx}{\ECMV}{\TMV}$ for synthesis and $\ctxAnaTypeM{\ctx}{\ECMV}{\TMV}$ for analysis.
Finally, marking is bidirectional as well. The synthetic marking judgment
$\ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$ establishes that, under the context $\Gamma$, the
unmarked expression $\EMV$ is ``marked into'' the marked expression $\ECMV$, which synthesizes type
$\TMV$. Analogously, the analytic marking judgment $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$
states that $\EMV$ is marked into $\ECMV$, which analyzes against $\TMV$.

How can we ensure that the marking procedure is correctly defined? There are two critical
metatheorems to keep in mind as we continue. The first is a \emph{totality} of marking:
%
\begin{theorem}[name=Marking Totality] \
  \label{thm:calculus-marking-totality}
  \begin{enumerate}
    \item For all $\ctx$ and $\EMV$, there exist $\ECMV$ and $\TMV$ such that
      $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$.
    \item For all $\ctx$, $\EMV$, and $\TMV$, there exists $\ECMV$ such that
      $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$.
  \end{enumerate}
\end{theorem}
%
That is, we may mark \emph{any} syntactically well-formed program.

\input{figs/calculus-mark-erasure}

Furthermore, since error marks are effectively annotations on top of the program, marking should
preserve syntactic structure modulo those marks. \cref{fig:calculus-mark-erasure} thus gives a
notion of \emph{mark erasure}, which converts marked expressions back into unmarked expression by
removing error marks. Then, \cref{thm:calculus-marking-well-formedness} provides a
\emph{well-formedness} criterion for marking.
%
\begin{theorem}[name=Marking Well-Formedness] \
  \label{thm:calculus-marking-well-formedness}
  \begin{enumerate}
    \item If $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$, then $\ctxSynTypeM{\ctx}{\ECMV}{\TMV}$ and
      $\erasesTo{\ECMV}{\EMV}$.
    \item If $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$, then $\ctxAnaTypeM{\ctx}{\ECMV}{\TMV}$ and
      $\erasesTo{\ECMV}{\EMV}$.
  \end{enumerate}
\end{theorem}
%
In other words, marking and then erasing error marks from the result gives the original unmarked
expression. We also ensure that marking gives well-typed marked expressions; indeed, their types
should match those synthesized or analyzed against in marking.

In the remainder of this subsection, we consider the different type errors that may arise in the
unmarked language while keeping in mind these two theorems; in this way, the traditional typing
semantics of the unmarked language guide the construction of the marking process.

\subsubsection{Numbers}
\label{sec:calculus-numbers}

To start, let us consider the simple case of numbers. Because of \emph{subsumption}, which is
discussed in more detail next, we need only define a synthesis rule for unmarked numbers, in which
number literals synthesize the type $\TNum$:
%
\begin{mathpar}
  \judgment{ }{
    \ctxSynTypeU{\ctx}{\ENumMV}{\TNum}
  }{USNum}
\end{mathpar}

How should numbers be marked? Quite straightforwardly, we simply give the same number as a marked
expression, synthesizing again $\TNum$. No type errors may occur in a single number, so we only need
the following rules for marking numbers and typing the results: 
%
\begin{mathpar}
  \judgment{
  }{
    \ctxSynFixedInto{\ctx}{\ENumMV}{\ECNumMV}{\TNum}
  }{MKSNum}

  \judgment{ }{
    \ctxSynTypeM{\ctx}{\ECNumMV}{\TNum}
  }{MSNum}
\end{mathpar}

For addition expressions, the type of both operands should be $\TNum$. To denote this in a
bidirectional system, they are analyzed against $\TNum$, giving the following typing rule for
unmarked addition expressions:
%
\begin{mathpar}
  \judgment{
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TNum} \\
    \ctxAnaTypeU{\ctx}{\EMV_2}{\TNum}
  }{
    \ctxSynTypeU{\ctx}{\EPlus{\EMV_1}{\EMV_2}}{\TNum}
  }{USPlus}
\end{mathpar}

The marking rule mirrors \textsc{USPlus} closely. Since an expected type for each operand is known,
we shift the responsibility for any type errors to them. Hence, we recursively mark each operand in
analytic mode and rebuild the marked addition expression. The typing rule for marked addition
expressions then mirrors \textsc{USPlus} exactly.
%
\begin{mathpar}
  \judgment{
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TNum} \\
    \ctxAnaFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TNum}
  }{
    \ctxSynFixedInto{\ctx}{\EPlus{\EMV_1}{\EMV_2}}{\ECPlus{\ECMV_1}{\ECMV_2}}{\TNum}
  }{MKSPlus}

  \judgment{
    \ctxAnaTypeM{\ctx}{\ECMV_1}{\TNum} \\
    \ctxAnaTypeM{\ctx}{\ECMV_2}{\TNum}
  }{
    \ctxSynTypeM{\ctx}{\ECPlus{\ECMV_1}{\ECMV_2}}{\TNum}
  }{MSPlus}
\end{mathpar}

\subsubsection{Subsumption}
\label{sec:calculus-subsumption}

Only synthetic rules were necessary for typing numbers and addition because of \emph{subsumption},
given below, which states that if an expression synthesizes a type, it may also be analyzed against
that type or any \emph{consistent} type.
\[%
  \judgment{
    \ctxSynTypeU{\ctx}{\EMV}{\TMV'} \\
    \consistent{\TMV}{\TMV'} \\
    \subsumable{\EMV}
  }{
    \ctxAnaTypeU{\ctx}{\EMV}{\TMV}
  }{UASubsume}
\]%
We rely on the notion of \emph{type consistency} from gradual type theory, which defines a reflexive
and symmetric (but not transitive) relation between types, writing $\consistent{\TMV_1}{\TMV_2}$ to
mean that $\TMV_1$ is consistent with $\TMV_2$. Defined in \cref{fig:calculus-consistency}, this
replaces the notion of \emph{type equality} to relate the unknown type to all other types.

\begin{figure}[htbp]
  \raggedright
  \judgbox{\ensuremath{\consistent{\TMV_1}{\TMV_2}}} $\TMV_1$ is consistent with $\TMV_2$
  %
  \begin{mathpar}
    \judgment{ }{
      \consistent{\TUnknown}{\TMV}
    }{TCUnknown1}

    \judgment{ }{
      \consistent{\TMV}{\TUnknown}
    }{TCUnknown2}

    \judgment{ }{
      \consistent{\TMV}{\TMV}
    }{TCRefl}

    \judgment{
      \consistent{\TMV_1}{\TMV_1'} \\
      \consistent{\TMV_2}{\TMV_2'} \\
    }{
      \consistent{\TArrow{\TMV_1}{\TMV_2}}{\TArrow{\TMV_1'}{\TMV_2'}}
    }{TCArr} \\
  \end{mathpar}
  %
  \caption{Type consistency.}
  \label{fig:calculus-consistency}
\end{figure}

Hence, when checking $\ENumMV$ against a known expected type, subsumption checks that the type is
consistent with $\TNum$, the type that $\ENumMV$ synthesizes. This succeeds for $\TUnknown$ and
$\TNum$.

We also restrict the usage of subsumption to ``subsumable'' syntactic forms, written
$\subsumable{\EMV}$. This judgment is defined on all syntactic forms except lambda abstractions and
conditionals, the only ones with explicit analysis rules. In other words, we restrict subsumption to
be the rule of ``last resort'', which is necessary to establish that marking and typing are
deterministic (discussed later in \cref{thm:calculus-marking-unicity}).

Now, to define analytic marking on forms without explicit analytic typing rules, we also need
subsumption rules for marking and marked expression typing. Note that we define a analogous notion
of ``subsumability'' for marked expressions, written $\subsumable{\ECMV}$.
%
\begin{mathpar}
  \judgment{
    \ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV'} \\
    \consistent{\TMV}{\TMV'} \\
    \subsumable{\EMV}
  }{
    \ctxAnaFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}
  }{MKASubsume}

  \judgment{
    \ctxSynTypeM{\ctx}{\ECMV}{\TMV'} \\
    \consistent{\TMV}{\TMV'} \\
    \subsumable{\ECMV}
  }{
    \ctxAnaTypeM{\ctx}{\ECMV}{\TMV}
  }{MASubsume}
\end{mathpar}

But what happens when the synthesized type of an expression is \emph{not} consistent with the
expected type, i.e. that the premise $\consistent{\TMV}{\TMV'}$ fails? Recalling the example of
\cref{fig:calculus-examples-inconsistent-types}, subsumption would be used to analyze $\ETrue$
against $\TNum$ in $\EPlus{\ETrue}{5}$, which fails since $\inconsistent{\TNum}{\TBool}$, i.e.
\textsc{UASubsume} does not apply, and traditional typing semantics would simply fail the type
checking process at this point. \textsc{MKASubsume} would also not be applicable, leaving marking
undefined in those cases.

% TODO use example more closely related to num here since we haven't talked about booleans yet?

To satisfy marking totality, such a possibility motivates an \emph{inconsistent type mark}
$\ECInconType{\ECMV}$, which states exactly that the synthesized type of $\ECMV$ is inconsistent
with the expected type. It is then accompanied by the following rules:
%
\begin{mathpar}
  \judgment{
    \ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV'} \\
    \inconsistent{\TMV}{\TMV'} \\
    \subsumable{\EMV}
  }{
    \ctxAnaFixedInto{\ctx}{\EMV}{\ECInconType{\ECMV}}{\TMV}
  }{MKAInconsistentTypes}

  \judgment{
    \ctxSynTypeM{\ctx}{\ECMV}{\TMV'} \\
    \inconsistent{\TMV}{\TMV'} \\
    \subsumable{\ECMV}
  }{
    \ctxAnaTypeM{\ctx}{\ECInconType{\ECMV}}{\TMV}
  }{MAInconsistentTypes}
\end{mathpar}
%
Observe that the premises of \textsc{MKAInconsistentTypes} are identical to those of
\textsc{MKASubsume}, except that $\inconsistent{\TMV}{\TMV'}$. By marking an error, type-checking
may continue---instead of simply failing.

\subsubsection{Variables}
\label{sec:calculus-variables}

Let us now consider the case of variables. Typing in the unmarked language is standard, and because
of subsumption, only a synthetic rule is required:
\[%
  \judgment{
    \inCtx{\ctx}{x}{\TMV}
  }{
    \ctxSynTypeU{\ctx}{x}{\TMV}
  }{USVar}
\]%
That is, if $x$ is bound to a type in the typing context, it synthesizes that type.
Straightforwardly, marking converts an unmarked variable into the same variable via the following
rules:
%
\begin{mathpar}
  \judgment{
    \inCtx{\ctx}{x}{\TMV}
  }{
    \ctxSynFixedInto{\ctx}{x}{x}{\TMV}
  }{MKSVar}

  \judgment{
    \inCtx{\ctx}{x}{\TMV}
  }{
    \ctxSynTypeM{\ctx}{x}{\TMV}
  }{MSVar}
\end{mathpar}
% TODO more intuition on why this makes sense?

However, consider the case that a variable, such as $y$ in \cref{fig:calculus-examples-unbound}, is \emph{not}
bound. Again, \textsc{USVar} would not apply. A total marking should, however, report the error and
continue, motivating an \emph{unbound variable mark} $\ECUnbound{x}$ and the accompanying rules:
%
\begin{mathpar}
  \judgment{
    \notInCtx{\ctx}{x}
  }{
    \ctxSynFixedInto{\ctx}{x}{\ECUnbound{x}}{\TUnknown}
  }{MKSUnbound}

  \judgment{
    \notInCtx{\ctx}{x}
  }{
    \ctxSynTypeM{\ctx}{\ECUnbound{x}}{\TUnknown}
  }{MSUnbound}
\end{mathpar}
%
An unbound variable is marked as such, and since nothing may be said about their types, we
synthesize the unknown type. Similarly to the inconsistent type mark, this allows type checking to
proceed, with usage of the unbound variable permitted in any expression.
% TODO clarify/give intuition as to why this is correct

\subsubsection{Lambda Abstractions}
\label{sec:calculus-lambda-abstractions}

Unlike numbers and variables, there are explicit synthesis and analysis rules for unmarked lambda
abstractions: % TODO intuition why
%
\begin{mathpar}
  \judgment{
    \ctxSynTypeU{\extendCtx{\ctx}{x}{\TMV}}{\EMV}{\TMV_2}
  }{
    \ctxSynTypeU{\ctx}{\ELam{x}{\TMV_1}{\EMV}}{\TArrow{\TMV_1}{\TMV_2}}
  }{USLam}

  \judgment{
    \matchedArrow{\TMV_3}{\TMV_1}{\TMV_2} \\
    \consistent{\TMV}{\TMV_1} \\
    \ctxAnaTypeU{\extendCtx{\ctx}{x}{\TMV}}{\EMV}{\TMV_2}
  }{
    \ctxAnaTypeU{\ctx}{\ECLam{x}{\TMV}{\EMV}}{\TMV_3}
  }{UALam}
\end{mathpar}
%
The synthesis rule is standard. The analysis rule employs the judgment
$\matchedArrow{\TMV}{\TMV_1}{\TMV_2}$, which establishes that $\TMV$ is a \emph{matched arrow type}
\cite{cimini:2016}, i.e. it may be considered an arrow type. Defined in
\cref{fig:calculus-matched-arrow}, this notion is purely a technical mechanism to avoid duplication
of rules related to arrow types.

% TODO mention explicitly that by inductive hypothesis, the marking of body cannot fail?
% TODO more intuition on these two judgments
% TODO more intuition on USLam and UALam

\begin{figure}[htbp]
  \raggedright
  \judgbox{\ensuremath{\matchedArrow{\TMV}{\TMV_1}{\TMV_2}}} $\TMV$ has matched arrow type $\TArrow{\TMV_1}{\TMV_2}$
  %
  \begin{mathpar}
    \judgment{ }{
      \matchedArrow{\TUnknown}{\TUnknown}{\TUnknown}
    }{TMAUnknown}

    \judgment{ }{
      \matchedArrow{\TArrow{\TMV}{\TMV}}{\TMV}{\TMV}
    }{TMAArr}
  \end{mathpar}
  %
  \caption{Matched arrow types.}
  \label{fig:calculus-matched-arrow}
\end{figure}

The synthesis rule for marking intuitively follows \textsc{USLam} closely. We should recursively
mark the body with an extended context binding $x$ to $\TMV$ and construct a new marked lambda
abstraction:
%
\begin{mathpar}
  \judgment{
    \ctxSynFixedInto{\extendCtx{\ctx}{x}{\TMV_1}}{\EMV}{\ECMV}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\ELam{x}{\TMV_1}{\EMV}}{\ELam{x}{\TMV_1}{\ECMV}}{\TArrow{\TMV_1}{\TMV_2}}
  }{MKSLam}

  \judgment{
    \ctxSynTypeM{\extendCtx{\ctx}{x}{\TMV}}{\ECMV}{\TMV_2}
  }{
    \ctxSynTypeM{\ctx}{\ECLam{x}{\TMV_1}{\ECMV}}{\TArrow{\TMV_1}{\TMV_2}}
  }{MSLam}
\end{mathpar}
%
Similarly, we construct corresponding analytic rules:
%
\begin{mathpar}
  \judgment{
    \matchedArrow{\TMV_3}{\TMV_1}{\TMV_2} \\
    \consistent{\TMV}{\TMV_1} \\\\
    \ctxAnaFixedInto{\extendCtx{\ctx}{x}{\TMV}}{\EMV}{\ECMV}{\TMV_2}
  }{
    \ctxAnaFixedInto{\ctx}{\ELam{x}{\TMV}{\EMV}}{\ECLam{x}{\TMV}{\ECMV}}{\TMV_3}
  }{MKALam1}

  \judgment{
    \matchedArrow{\TMV_3}{\TMV_1}{\TMV_2} \\
    \consistent{\TMV}{\TMV_1} \\\\
    \ctxAnaTypeM{\extendCtx{\ctx}{x}{\TMV}}{\ECMV}{\TMV_2}
  }{
    \ctxAnaTypeM{\ctx}{\ECLam{x}{\TMV}{\ECMV}}{\TMV_3}
  }{MALam1}
\end{mathpar}
%
However, recalling totality, we look again to the premises of \textsc{UALam} to see what type errors
may arise. First, what if $\TMV_3$ is not a matched arrow type? This would be the case if it were
$\TNum$, for example. The lambda abstraction's synthesized type would indeed be inconsistent with
the expected type, but this case is slightly distinct from that of the inconsistent types mark: the
lambda expression is in analytic position. Instead, we mark $\ECAnaNonMatchedArrow{\ECMV}$ to
indicate that an expression of the non-matched arrow type $\TMV_3$ was expected, but a lambda
abstraction was encountered.
%
\begin{mathpar}
  \judgment{
    \notMatchedArrow{\TMV_3} \\
    \ctxAnaFixedInto{\extendCtx{\ctx}{x}{\TMV}}{\EMV}{\ECMV}{\TUnknown}
  }{
    \ctxAnaFixedInto{\ctx}{\ELam{x}{\TMV}{\EMV}}{\ECLamAnaNonMatchedArrow{x}{\TMV}{\ECMV}}{\TMV_3}
  }{MKALam2}

  \judgment{
    \notMatchedArrow{\TMV_3} \\
    \ctxAnaTypeM{\extendCtx{\ctx}{x}{\TMV}}{\ECMV}{\TUnknown}
  }{
    \ctxAnaTypeM{\ctx}{\ECLamAnaNonMatchedArrow{x}{\TMV}{\ECMV}}{\TMV_3}
  }{MALam2}
\end{mathpar}
%
Though no expected output type is known, we still need to check and mark the body; it is thus
analyzed against the unknown type in \textsc{MKALam2}.

In this example language, the distinction from the inconsistent type mark is of reduced
significance---all expressions synthesize a type. However, the addition of unannotated lambda
abstractions, for example, would necessitate such a disctinction.

Another error arises when $\matchedArrow{\TMV_3}{\TMV_1}{\TMV_2}$, but
$\inconsistent{\TMV}{\TMV_1}$, i.e. the actual type annotation for $x$ is inconsistent with the
expected input type. The \emph{inconsistent ascription mark} $\ECInconAsc{x}$ indicates exactly this
error, and we add a final pair of analytic rules:
%
\begin{mathpar}
  \judgment{
    \matchedArrow{\TMV_3}{\TMV_1}{\TMV_2} \\
    \inconsistent{\TMV}{\TMV_1} \\
    \ctxAnaFixedInto{\extendCtx{\ctx}{x}{\TMV_1}}{\EMV}{\ECMV}{\TMV_2}
  }{
    \ctxAnaFixedInto{\ctx}{\ELam{x}{\TMV}{\EMV}}{\ECLamInconAsc{x}{\TMV}{\ECMV}}{\TMV_3}
  }{MKALam3}

  \judgment{
    \matchedArrow{\TMV_3}{\TMV_1}{\TMV_2} \\
    \inconsistent{\TMV}{\TMV_1} \\
    \ctxAnaTypeM{\extendCtx{\ctx}{x}{\TMV_1}}{\ECMV}{\TMV_2}
  }{
    \ctxAnaTypeM{\ctx}{\ECLamInconAsc{x}{\TMV}{\ECMV}}{\TMV_3}
  }{MALam3}
\end{mathpar}

\subsubsection{Applications}
\label{sec:calculus-applications}

In the unmarked language, only a synthesis rule is necessary for applications:
\[%
  \judgment{
    \ctxSynTypeU{\ctx}{\EMV_1}{\TMV} \\
    \matchedArrow{\TMV}{\TMV_1}{\TMV_2} \\
    \ctxAnaTypeU{\ctx}{\EMV_2}{\TMV_1}
  }{
    \ctxSynTypeU{\ctx}{\EAp{\EMV_1}{\EMV_2}}{\TMV_2}
  }{USAp}
\]%
Following the same methodology, we can straightforwardly give the following marking and typing
rules:
%
\begin{mathpar}
  \judgment{
    \ctxSynFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TMV} \\
    \matchedArrow{\TMV}{\TMV_1}{\TMV_2} \\\\
    \ctxAnaFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV_1} \\
  }{
    \ctxSynFixedInto{\ctx}{\EAp{\EMV_1}{\EMV_2}}{\ECAp{\ECMV_1}{\ECMV_2}}{\TMV_2}
  }{MKSAp1}

  \judgment{
    \ctxSynTypeM{\ctx}{\ECMV_1}{\TMV} \\
    \matchedArrow{\TMV}{\TMV_1}{\TMV_2} \\\\
    \ctxAnaTypeM{\ctx}{\ECMV_2}{\TMV_1}
  }{
    \ctxSynTypeM{\ctx}{\ECAp{\ECMV_1}{\ECMV_2}}{\TMV_2}
  }{MSAp1}
\end{mathpar}

Again, to satisfy totality, we must consider the case when $\TMV$ is not a matched arrow type, such
as in the example of \cref{fig:calculus-examples-app-non-lambda}. There is no expected type for the
argument, so we perform analytic marking on $\EMV_2$ against the unknown type. In this case, it is
not quite right to mark $\EMV_1$ with the inconsistent type mark; rather than any single type, it is
any type within the family of arrow types that is expected. For such a ``constrained'' synthetic
mode, we use a specialized mark $\ECSynNonMatchedArrow{\ECMV}$, which indicates that $\ECMV$ was
expected to be a function---but was not. The output type is unknown, so the entire expression
synthesizes the unknown type.
%
\begin{mathpar}
  \judgment{
    \ctxSynFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TMV} \\
    \notMatchedArrow{\TMV} \\
    \ctxAnaFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TUnknown}
  }{
    \ctxSynFixedInto{\ctx}{\EAp{\EMV_1}{\EMV_2}}{\ECApSynNonMatchedArrow{\ECMV_1}{\ECMV_2}}{\TUnknown}
  }{MKSAp2}

  \judgment{
    \ctxSynTypeM{\ctx}{\ECMV_1}{\TMV} \\
    \notMatchedArrow{\TMV} \\
    \ctxAnaTypeM{\ctx}{\ECMV_2}{\TUnknown}
  }{
    \ctxSynTypeM{\ctx}{\ECApSynNonMatchedArrow{\ECMV}{\ECMV}}{\TUnknown}
  }{MSAp2}
\end{mathpar}

It is natural to extend this approach to other elimination forms that require the handling of
unmatched types, such as products. Indeed, the same approach is taken for projections in
\cref{sec:calculus-products}. Another approach might indicate the same kind of error but mark the
entire application.

\subsubsection{Booleans}
\label{sec:calculus-booleans}

The boolean values are straightforward to handle, behaving in a similar manner as numbers: 
%
\begin{mathpar}
  \judgment{ }{
    \ctxSynTypeU{\ctx}{\ETrue}{\TBool}
  }{USTrue}

  \judgment{
  }{
    \ctxSynFixedInto{\ctx}{\ETrue}{\ECTrue}{\TBool}
  }{MKSTrue}

  \judgment{ }{
    \ctxSynTypeM{\ctx}{\ECTrue}{\TBool}
  }{MSTrue} \\

  \judgment{ }{
    \ctxSynTypeU{\ctx}{\EFalse}{\TBool}
  }{USFalse}

  \judgment{
  }{
    \ctxSynFixedInto{\ctx}{\EFalse}{\ECFalse}{\TBool}
  }{MKSFalse}

  \judgment{ }{
    \ctxSynTypeM{\ctx}{\ECFalse}{\TBool}
  }{MSFalse}
\end{mathpar}

Conditionals, however, offer an interesting case. In the unmarked language, we have both explicit
synthetic and analytic rules:
%
\begin{mathpar}
  \judgment{
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TBool} \\\\
    \ctxSynTypeU{\ctx}{\EMV_2}{\TMV_1} \\
    \ctxSynTypeU{\ctx}{\EMV_3}{\TMV_2}
  }{
    \ctxSynTypeU{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\TJoin{\TMV_1}{\TMV_2}}
  }{USIf}

  \judgment{
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TBool} \\\\
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TMV} \\
    \ctxAnaTypeU{\ctx}{\EMV_2}{\TMV}
  }{
    \ctxAnaTypeU{\ctx}{\ECIf{\EMV_1}{\EMV_2}{\EMV_3}}{\TMV}
  }{UAIf}
\end{mathpar}
%
In synthetic position, conditionals synthesize the \emph{join} of the branch types $\TMV_1$ and
$\TMV_2$, which we define inductively in \cref{fig:calculus-type-join}. In analytic position, since
there is an expected type for both branches, we shift the blame for any errors to them.

% TODO intuition why we use join

\newcommand{\joinsTo}[3]{\ensuremath{\TJoin{#1}{#2} & = & #3}}
\begin{figure}[htbp]
  \raggedright
  \judgbox{\ensuremath{\TJoin{\TMV_1}{\TMV_2}}} is a \emph{partial} metafunction defined as follows:
  %
  \[\begin{array}{rcl}
    \joinsTo{\TUnknown}{\TMV}{\TMV} \\
    \joinsTo{\TMV}{\TUnknown}{\TMV} \\
    \joinsTo{\TNum}{\TNum}{\TNum} \\
    \joinsTo{\TBool}{\TBool}{\TBool} \\
    \joinsTo{(\TArrow{\TMV_1}{\TMV_2})}{(\TArrow{\TMV_1'}{\TMV_2'})}{\TArrow{(\TJoin{\TMV_1}{\TMV_1'})}{(\TJoin{\TMV_2}{\TMV_2'})}} \\
  \end{array}\]
  %
  \caption{Type join.}
  \label{fig:calculus-type-join}
\end{figure}
% TODO intuition for type join?

Following \textsc{USIf} and \textsc{UAIf}, we may derive the following marking and typing rules.
Similar to previous cases, we recursively mark the guard and branches, constructing the marked
conditional from the results.
%
\begin{mathpar}
  \judgment{
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TBool} \\\\
    \ctxSynFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV_1} \\
    \ctxSynFixedInto{\ctx}{\EMV_3}{\ECMV_3}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\ECIf{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TJoin{\TMV_1}{\TMV_2}}
  }{MKSIf}

  \judgment{
    \ctxAnaTypeM{\ctx}{\ECMV_1}{\TBool} \\\\
    \ctxSynTypeM{\ctx}{\ECMV_2}{\TMV_1} \\
    \ctxSynTypeM{\ctx}{\ECMV_3}{\TMV_2}
  }{
    \ctxSynTypeM{\ctx}{\ECIf{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TJoin{\TMV_1}{\TMV_2}}
  }{MSIf} \\

  \judgment{
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TBool} \\\\
    \ctxAnaFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV} \\
    \ctxAnaFixedInto{\ctx}{\EMV_3}{\ECMV_3}{\TMV} \\
  }{
    \ctxAnaFixedInto{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\ECIf{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TMV}
  }{MKAIf}

  \judgment{
    \ctxAnaTypeM{\ctx}{\ECMV_1}{\TBool} \\\\
    \ctxAnaTypeM{\ctx}{\ECMV_1}{\TMV} \\
    \ctxAnaTypeM{\ctx}{\ECMV_2}{\TMV}
  }{
    \ctxAnaTypeM{\ctx}{\ECIf{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TMV}
  }{MAIf}
\end{mathpar}

However, in synthetic position, it may be the case that the two branch types have no join. This
occurs, in fact, when they are inconsistent, motivating the \emph{the inconsistent branches mark}
$\ECInconBr{\ECMV}{\ECMV}{\ECMV}$. Then, adding the following rules ensures totality on
conditionals:
%
\begin{mathpar}
  \judgment{
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TBool} \\
    \ctxSynFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV_1} \\\\
    \ctxSynFixedInto{\ctx}{\EMV_3}{\ECMV_3}{\TMV_2} \\
    \inconsistent{\TMV_1}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\ECInconBr{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TUnknown}
  }{MKSInconsistentBranches}

  \judgment{
    \ctxAnaTypeM{\ctx}{\ECMV_1}{\TBool} \\
    \ctxSynTypeM{\ctx}{\ECMV_2}{\TMV_1} \\\\
    \ctxSynTypeM{\ctx}{\ECMV_3}{\TMV_2} \\
    \inconsistent{\TMV_1}{\TMV_2}
  }{
    \ctxSynTypeM{\ctx}{\ECInconBr{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TUnknown}
  }{MSInconsistentBranches}
\end{mathpar}

The framework also allows for other localization approaches. For example, as discussed at the
beginning of the section, we may regard the first branch as ``correct'' and localize any errors to
the second. The following rule formalizes such a design:
\[%
  \judgment{
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TBool} \\
    \ctxSynFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV} \\
    \ctxAnaFixedInto{\ctx}{\EMV_3}{\ECMV_3}{\TMV}
  }{
    \ctxSynFixedInto{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\ECIf{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TMV}
  }{MKSIf'}
\]%
Of course, corresponding unmarked and marked typing rules would follow similar structure, and other
heuristics may be devised for different kinds of branching structures.

\subsubsection{Holes}
For completeness, we finish the development of the marked lambda calculus on the extended GTLC with
the marking of holes. Like number literals and booleans, these are governed by straightforward
rules.
%
\begin{mathpar}
  \judgment{ }{
    \ctxSynTypeU{\ctx}{\EEHole}{\TUnknown}
  }{USHole}

  \judgment{ }{
    \ctxSynFixedInto{\ctx}{\EEHole}{\ECEHole}{\TUnknown}
  }{MKSHole}

  \judgment{ }{
    \ctxSynTypeM{\ctx}{\ECEHole}{\TUnknown}
  }{MSHole}
\end{mathpar}

% TODO mini conclusion on methodology

\subsubsection{Additional Theorems}
To conclude this section, we present two more metatheorems that help ensure correctness of the
system. Though \cref{thm:calculus-marking-well-formedness} guarantees that marking does not change
the syntactic structure of a program, it makes no statement about the presence of error marks in the
resulting marked program. Following the logic of constructing marking rules, we have
\cref{thm:calculus-marking-well-ill-typed}.

\begin{theorem}[name=Marking of Well-Typed/Ill-Typed Expressions] \ % TODO better name for this
  \label{thm:calculus-marking-well-ill-typed}
  \begin{enumerate}
    \item \begin{enumerate}
        \item If $\ctxSynTypeU{\ctx}{\EMV}{\TMV}$ and $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$,
          then $\markless{\ECMV}$.
        \item If $\ctxAnaTypeU{\ctx}{\EMV}{\TMV}$ and $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV}{\TMV}$,
          then $\markless{\ECMV}$.
      \end{enumerate}

    \item \begin{enumerate}
        \item If there does not exist $\TMV$ such that $\ctxSynTypeU{\ctx}{\EMV}{\TMV}$, then for
          all $\ECMV$ and $\TMV'$ such that $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV'}$, it is not
          the case that $\markless{\ECMV}$.
        \item If there does not exist $\TMV$ such that $\ctxAnaTypeU{\ctx}{\EMV}{\TMV}$, then for
          all $\ECMV$ and $\TMV'$ such that $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV}{\TMV'}$, it is not
          the case that $\markless{\ECMV}$.
      \end{enumerate}
  \end{enumerate}
\end{theorem}
%
That is, since error marks indicate type errors in the program, no marks should be introduced
when marking a well-typed program, a property given by the judgment $\markless{\ECMV}$. The
converse---that at least one mark is inserted into ill-typed programs---is also true.

Finally, no less importantly, marking is deterministic. This is given by
\cref{thm:calculus-marking-unicity}.
%
\begin{theorem}[name=Marking Unicity] \
  \label{thm:calculus-marking-unicity}
  \begin{enumerate}
    \item If $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV_1}{\TMV_1}$ and
      $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV_2}{\TMV_2}$, then $\ECMV_1 = \ECMV_2$ and $\TMV_1 =
      \TMV_2$.
    \item If $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV_1}{\TMV}$ and
      $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV_2}{\TMV}$, then $\ECMV_1 = \ECMV_2$.
  \end{enumerate}
\end{theorem}
%
Together, totality and unicity give that marking may be considered and implemented as a total
function. Indeed, given the algorithmic nature of bidirectional typing, it is fairly easy to extract
an implementation from the rules.

\subsection{Adding Binary Products}
\label{sec:calculus-products}

We now demonstrate how one might extend or adapt the marked lambda calculus for new typing features
by adding binary product types, written $\TProd{\TMV}{\TMV}$. In the interest of space, we focus on
aspects related to marking; a comprehensive set of rules is found in the supplemental material.

First, we extend the syntax of types and the accompanying notions of consistency and join, which
closely mirror the definitions for arrow types. Additionally, analogous to the notion of matched
arrow types, the judgment $\matchedProd{\TMV}{\TMV_1}{\TMV_2}$ establishes that $\TMV$ may be
considered the binary product type $\TProd{\TMV_1}{\TMV_2}$ [cite?].

The unmarked language is extended with a pair constructor $\EPair{\EMV}{\EMV}$ and projection
operators $\EProjL{\EMV}$ and $\EProjR{\EMV}$, governed by standard bidirectional rules. In addition
to the synthesis rules, an explicit analysis rules is defined for the former: if an expected type is
known, the responsibility for type errors is shift to the components.

The marked language is extended with the same forms, and we again devise marking rules by modelling
off the unmarked typing rules.
%
\begin{mathpar}
  \judgment{
    \ctxSynFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TMV_1} \\\\
    \ctxSynFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\EPair{\EMV_1}{\EMV_2}}{\ECPair{\ECMV_1}{\ECMV_2}}{\TProd{\TMV_1}{\TMV_2}}
  }{MKSPair}

  \judgment{
    \ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV} \\\\
    \matchedProd{\TMV}{\TMV_1}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\EProjL{\EMV}}{\ECProjL{\ECMV}}{\TMV_1}
  }{MKSProjL1}

  \judgment{
    \ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV} \\\\
    \matchedProd{\TMV}{\TMV_1}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\EProjR{\EMV}}{\ECProjR{\ECMV}}{\TMV_2}
  }{MKSProjR1} \\

  \judgment{
    \matchedProd{\TMV}{\TMV_1}{\TMV_2} \\
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TMV_1} \\\\
    \ctxAnaFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV_2}
  }{
    \ctxAnaFixedInto{\ctx}{\EPair{\EMV_1}{\EMV_2}}{\ECPair{\ECMV_1}{\ECMV_2}}{\TMV}
  }{MKAPair1}
\end{mathpar}

However, as previously seen with lambda abstractions and applications, two sets of rules are further
necessary to satisfy totality. Firstly, the premise in \textsc{UAPair} that $\TMV$ is a matched
binary product may fail, requiring the introduction of a new error mark,
$\ECAnaNonMatchedProd{\ECMV}$, which is analogous to $\ECAnaNonMatchedArrow{\ECMV}$ for lambda
abstractions. Another kind of error may arise when the subject of a projection does not synthesize a
matched product type. As in our development of marking for applications, we utilize a new mark,
written $\ECSynNonMatchedProd{\ECMV}$. These cases give the final marking rules for products:
%
\begin{mathpar}
  \judgment{
    \notMatchedProd{\TMV} \\
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TUnknown} \\
    \ctxAnaFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TUnknown}
  }{
    \ctxAnaFixedInto{\ctx}{\EPair{\EMV_1}{\EMV_2}}{\ECPairAnaNonMatchedProd{\ECMV_1}{\ECMV_2}}{\TMV}
  }{MKAPair2} \\

  \judgment{
    \ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV} \\
    \notMatchedProd{\TMV}
  }{
    \ctxSynFixedInto{\ctx}{\EProjL{\EMV}}{\ECProjLSynNonMatchedProd{\ECMV}}{\TMV_1}
  }{MKSProjL2}

  \judgment{
    \ctxSynFixedInto{\ctx}{\EMV}{\ECMV}{\TMV} \\
    \notMatchedProd{\TMV}
  }{
    \ctxSynFixedInto{\ctx}{\EProjR{\EMV}}{\ECProjRSynNonMatchedProd{\ECMV}}{\TMV_2}
  }{MKSProjR2}
\end{mathpar}

\input{let}

\subsection{Polymorphism and Other Typing Features}
\label{sec:calculus-poly}

We now consider how the calculus might be extended to richer typing features such as parametric
polymorphism. Specifically, we consider System F-style explicit polymorphism.

\subsection{Mechanization}
\label{sec:calculus-agda}

The semantics and metatheory presented above, including the extension with binary products, have
been fully mechanized in the Agda proof assistant. Though the mechanization's documentation contains
more detailed discussion regarding technical decisions made therein, we highlight some important
aspects here.

The standard approach of modeling judgments as inductive datatypes and rules as constructors for
those datatypes is taken. For marked expressions, we opt for implicitly typed terms [citation],
which gives us part of \cref{thm:calculus-marking-well-formedness} for free. For the convenience of
readers interested in browsing the mechanization, all rule names match those presented here.

\section{The Marked Lambda in Structured Editing}
\label{sec:calculus-structured-editing}

[Something about structure editors]

The previous section presented a formal calculus for error localization and recovery. We now discuss
its usage in a structured editing setting. In particular, the calculus has been integrated into the
Hazel programming environment, a live programming environment with support for typed holes.

\subsection{Implementation in Hazel}
\label{sec:calculus-hazel}

\subsection{Reified Hazelnut Action Semantics}
\label{sec:calculus-hazelnut}

The Hazelnut action calculus introduced by \citet{HazelnutPOPL} gives a structure editing scheme
that employs \emph{expression holes}, denoted $\EEHole$, which represent syntactically incomplete
part of the program. Not only are terms always syntactically well-formed, they are always
\emph{well-typed}, ensuring that semantic services are always available. To accomplish this, the
bidirectional action calculus---given by the synthetic action judgment, written
$\ASEAction{\ctx}{\EMV}{\TMV}{\EMV'}{\TMV'}{\AMV}$, and the analytic action judgment, written
$\AAEAction{\ctx}{\EMV}{\EMV'}{\TMV}{\AMV}$---inserts on-demand \emph{error holes}, denoted
$\hhole{\EMV}{}$, which indicate type errors.

% TODO example

Hazelnut, however, is unable to cope with actions that require \emph{non-local} hole fixes. For
example, though the system can wrap an arbitrary expression in an arithmetic operation, it cannot
wrap an arbitrary expression $\EMV$ in a lambda abstraction, i.e. construct
$\ELam{x}{\TUnknown}{\EMV}$ directly from $\EMV$. The binding of $x$ with the unknown type may
shadow a previous declaration and require the removal of error holes anywhere in body, but Hazelnut
may only perform such fixes for immediately local holes. Such actions remain undefined, i.e.
unperformable.

By applying the marked lambda calculus to Hazelnut, the action calculus may be \emph{wholly decoupled}
from the typing semantics. Under this new system, the action calculus, given by the singular action
judgment, written $\AUEAction{\EMV}{\EMV'}{\AMV}$, is concerned only with the manipulation of
syntax, and the marking procedure yields statically meaningful terms with error marks, now error
holes, inserted at the appropriate positions as a kind of editor decoration. This solves the problem
outlined above; since marking operates on the entire program after each action, arbitrary
constructions are permitted.

Alternatively, instead of a wholly untyped action calculus, one may re-adapt the existing Hazelnut
calculus with scope-based marking. Action operate over \emph{marked expressions}, and error marks
operate as error holes. The system is once again bidirectional, writing (note $\ECMV$ instead of
$\EMV$) the judgments $\ASEAction{\ctx}{\ECMV}{\TMV}{\ECMV'}{\TMV'}{\AMV}$ and
$\AAEAction{\ctx}{\ECMV}{\ECMV'}{\TMV}{\AMV}$. Then, when wrapping $\ECMV$ into
$\ECLam{x}{\TUnknown}{\ECMV}$, we may simply re-mark the body under the context with $x$ binding
the unknown type:
\[%
  \judgment{
    \ctxSynFixedInto{\extendCtx{\ctx}{x}{\TUnknown}}{\erase{\ECMV}}{\ECMV'}{\TMV'}
  }{
    \ASEConLam{\ctx}{\ECMV}{\TMV}{\ECMV'}{\TMV'}{x}
  }{ASEConLam}
\]%
This system optimizes over the untyped version by performing marking only when necessary. In
particular, whenever binders are introduced around an expression, marking redraws error holes. The
supplemental material contains a complete formal description of both variants, and the untyped
action semantics and related metatheory are mechanized in Agda.

% TODO note that we have not verified the metatheory for the typed system