### 2023-03-02 conclusion.tex
Changes from Overleaf

\section{Conclusion}
\label{sec:conclusion}

Programming is increasingly a live collaboration between human programmers and sophisticated semantic services. 
These services need to be able to reason throughout the programming process, not just when the program is already complete. This paper lays down rigorous type-theoretic foundations for doing just that. Local type inference helps make the localization decisions we make systematic and predictable, and type hole inference shows how local and constraint-based type inference might operate hand-in-hand, rather than as alternatives. 
We hope that language designers will use the techniques introduced in this paper to consider more rigorously, perhaps even formally, the problems of type error localization and error recovery when designing future languages.