### 2023-07-11 polymorphism.tex

\subsection{Parametric Polymorphism}
\label{sec:calculus-polymorphism}

To demonstrate that the judgmental structure of the marked lambda calculus may be applied to richer typing features, we now explore an extension of the core language developed above to System F-style parametric polymorphism.

\[\begin{array}{rrcl}
  \TMName  & \TMV  & \Coloneqq & \cdots \mid \TForall{\TVarMV}{\TMV} \mid \TVarMV \\
  \MTMName & \MTMV & \Coloneqq & \cdots
                     % \MTUnknown \mid \MTNum \mid \MTBool
                     % \mid \MTArrow{\MTMV}{\MTMV} \mid \MTProd{\MTMV}{\MTMV}
                     \mid \MTForall{\MTVarMV}{\MTMV} \mid \MTVarMV \mid \MTUnbound{\MTVarMV} \\
  \EMName  & \EMV  & \Coloneqq & \cdots \mid \ETypeLam{\TVarMV}{\EMV} \mid \ETypeAp{\EMV}{\TMV} \\
  \ECMName & \ECMV & \Coloneqq & \cdots \mid \ECTypeLam{\MTVarMV}{\ECMV} \mid \ECTypeAp{\ECMV}{\MTMV} \\
           &       & \mid      & \ECSynNonMatchedForall{\ECMV} \mid \ECAnaNonMatchedForall{\ECMV}
\end{array}\]