### 2023-07-12 figs/calculus-examples.tex
Updates from Overleaf

\begin{figure}[htbp]
  \begin{tabular}[b]{cc}
    \begin{subfigure}[b]{0.3\columnwidth}
      \includegraphics[width=\columnwidth]{images/haz3l-unbound-variable.png}
      \caption{Free variable error.}
      \label{fig:calculus-examples-Free}
    \end{subfigure}
    &
    \begin{subfigure}[b]{0.3\columnwidth}
      \includegraphics[width=\columnwidth]{images/haz3l-inconsistent-types.png}
      \caption{Inconsistent types error.}
      \label{fig:calculus-examples-inconsistent-types}
    \end{subfigure} \\
    \begin{subfigure}[b]{0.3\columnwidth}
      \includegraphics[width=\columnwidth]{images/haz3l-ap-non-matched.png}
      \caption{Application of a non-lambda.}
      \label{fig:calculus-examples-app-non-lambda}
    \end{subfigure}
    &
    \begin{subfigure}[b]{0.3\columnwidth}
      \includegraphics[width=\columnwidth]{images/haz3l-inconsistent-branches.png}
      \caption{Inconsistent branches error.}
      \label{fig:calculus-examples-inconsistent-branches}
    \end{subfigure}
  \end{tabular}
  %
  \caption{Examples of common type errors.}
  \label{fig:calculus-examples}
\end{figure}