### 2023-07-12 supplement/preface.tex
Compile updated supplement

\documentclass[supplement.tex]{subfiles}

\begin{document}
\section{Preface}
\label{sec:preface}
This is a formalism demonstrating the \emph{marked lambda calculus}, a judgmental framework for
precise bidirectional error localization and recovery that employs gradual typing.

\subsection{Organization}
Though more is said in each individual section, the overall structure of the document is as follows:
%
\begin{itemize}
  \item \cref{sec:marked} employs the framework on a gradually typed lambda calculus.

  \item \cref{sec:patterned} extends the demonstration with patterned let expressions.

  \item \cref{sec:polymorphism} extends the demonstration with System F-style parametric
    polymorphism.

  \item \cref{sec:untyped} gives a version of the Hazelnut structure editor calculus that uses the
    marked lambda calculus to solve Hazelnut's deficiency with regards to non-local hole fixing.

  \item \cref{sec:typed} is similar, except that it employs the marking procedure in a roughly
    incremental fashion.
\end{itemize}
%
Note that each of the following sections after \cref{sec:marked} build upon the same core language.

\subsection{Mechanization}
Not all parts of the formalism are mechanized in Agda. In particular, the core of the marked lambda
calculus (\cref{sec:marked}) and the untyped Hazelnut calculus that employs marking
(\cref{sec:untyped}) \emph{are} mechanized, while all others \emph{are not}. Crucially, the
metatheorems of the aforementioned sections are verified, whereas those of all other sections are
not.

As possible, the names of judgments and rules that appear in the mechanization have been made to
follow those in this formalism. Refer also to the mechanization's README for more details.

\end{document}